---
title: "20221215_main_analysis"
author: "Yurun (Ellen) Ying"
date: "2022-12-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse);library(haven);library(here);library(broom);library(wesanderson)

# input data
bac_data <- 
  here("summative/data/hansen_dwi.dta") %>% 
  read_dta() %>% 
  mutate(
    # code into drive under influence (dui) = 1 if bac1 >= 0.08
    dui = ifelse(bac1 >= 0.08, 1, 0),
    
    # create a centered bac1 variable
    bac1_ctd = bac1 - 0.08,
    
    # create a quadratic term from bac1
    bac1_sq = bac1^2
    
    )

source(here("summative/code/summaryR.R"))
```

## Main analysis

Run the main analysis of the effect of treatment (punishment) on recidivism.

```{r}
# analysis officially given in the do file
# get the data within the 0.05 bandwidth
dat_bandwidth_0.05 <- bac_data %>% filter(abs(bac1_ctd) <= 0.05)

# linear regression
recidivism_linear_a <- 
  dat_bandwidth_0.05 %>% 
  lm(recidivism ~ dui*bac1 + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_linear_a, type = "hc1")

# quadratic model
recidivism_qua_a <- 
  dat_bandwidth_0.05 %>% 
  lm(recidivism ~ dui*(bac1 + bac1_sq) + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_qua_a, type = "hc1")


# get the data within the 0.025 bandwidth
dat_bandwidth_0.025 <- bac_data %>% filter(abs(bac1_ctd) <= 0.025)

# linear regression
recidivism_linear_b <- 
  dat_bandwidth_0.025 %>% 
  lm(recidivism ~ dui*bac1 + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_linear_b, type = "hc1")

# quadratic model
recidivism_qua_b <- 
  dat_bandwidth_0.025 %>% 
  lm(recidivism ~ dui*(bac1 + bac1_sq) + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_qua_b, type = "hc1")
```

The linear model estimates indicate that receiving punishment decreased the rate of recidivism by 5.91 percentage point (6.4 percentage when using 0.025 bandwidth). The quadratic model estimates indicate that receiving punishment decreased the rate of recidivism by 11.3 percentage point (37.1 percentage when using 0.025 bandwidth).

Make the scatter plots

```{r}
#min_bac <- min(dat_bandwidth_0.05$bac1) 
min_bac <- .030999999
max_bac_lower <- filter(dat_bandwidth_0.05, bac1_ctd < 0)$bac1 %>% max()
max_bac <- max(dat_bandwidth_0.05$bac1)
breaks <- c(seq(min_bac, max_bac_lower, length.out = 43), 
               seq(0.08, max_bac, length.out = 49)[-1])


dat_bandwidth_0.05 %>% 
  # create 90 bins with one break at the cutoff point
  mutate(bins = cut(bac1, breaks = breaks)) %>% 
   # select(bins, bac1) %>% arrange(bac1) %>% View()
  group_by(bins) %>% 
  summarize(
    recidivism = mean(recidivism), # calculate the mean recidivism in each group
    bac = mean(bac1), # roughly the midpoint of each bin
    dui = ifelse(bac >=0.08, 1, 0) %>% factor()
  ) %>% 
  ggplot(aes(x = bac, y = recidivism, color = dui)) +
  geom_point(show.legend = FALSE) +
  geom_vline(xintercept = 0.08, color = "red", linetype = 2) +
  scale_x_continuous(breaks = seq(0.02, 0.14, by = 0.02)) +
  scale_y_continuous(breaks = seq(0.07, 0.15, by = 0.02)) +
  scale_color_discrete(type = wes_palette("Royal1", 2, type = "discrete")) +
  labs(x = "Blood Alcohol Content (BAC)", y = "Recidivism") +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    panel.grid.major.y = element_line(color = "grey80", size = 0.4),
    axis.line = element_line(size = 0.4),
    axis.text.y = element_text(size = 11, margin = margin(0, 5, 0, 0)),
    axis.text.x = element_text(size = 11, margin = margin(5, 0, 0, 0)),
    axis.ticks = element_line(size = 0.4),
    axis.ticks.length = unit(6, units = "pt"),
    axis.title.y = element_text(size = 13, margin = margin(0, 10, 0, 0)),
    axis.title.x = element_text(size = 13, margin = margin(10, 0, 0, 0)),
    plot.margin = margin(20, 20, 20, 20)
  )
         
```


## Donut hole regression

Since there is heaping in our data, we will run a donut hole regression and check the robustness of our results.

```{r}
# add donut hole
dat_bandwidth_0.05 <- 
  # observation are dropped within one unit next to the cutoff point
  dat_bandwidth_0.05 %>% mutate(donut = ifelse(abs(bac1_ctd) <= 0.001, 1, 0))

# linear regression
recidivism_linear_a_donut <- 
  dat_bandwidth_0.05 %>% 
  filter(donut == 0) %>% 
  lm(recidivism ~ dui*bac1 + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_linear_a_donut, type = "hc1")

# narrower bandwidth
dat_bandwidth_0.025 <- 
  dat_bandwidth_0.025 %>% mutate(donut = ifelse(abs(bac1_ctd) <= 0.001, 1, 0))

# linear regression
recidivism_linear_b_donut <- 
  dat_bandwidth_0.025 %>% 
  filter(donut == 0) %>% 
  lm(recidivism ~ dui*bac1 + white + male + acc + aged,
     data = .)
summaryR.lm(recidivism_linear_b_donut, type = "hc1")
```

The donut hole regressions indicate that receiving punishment decreased the rate of recidivism by 5.57 percentage point (and 5.62 pp using 0.025 bandwidth).


## Notes for improvement

- Run local polynomials and explain why they are needed
- Consider what controls to put in the model and how they should be put
