---
title: "20221215_covariate_assumption_check"
author: "Yurun (Ellen) Ying"
date: "2022-12-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse);library(haven);library(here);library(broom)

# input data
bac_data <- 
  here("summative/data/hansen_dwi.dta") %>% 
  read_dta() %>% 
  mutate(
    # code into drive under influence (dui) = 1 if bac1 >= 0.08
    dui = ifelse(bac1 >= 0.08, 1, 0),
    
    # create a centered bac1 variable
    bac1_ctd = bac1 - 0.08,
    
    # create a quadratic term from bac1
    bac1_sq = bac1^2
    
    )

source(here("summative/code/summaryR.R"))
```

We will check the assumptions that observed covariates didn't have a jump at the cutoff point.

```{r}
# analysis officially given in the do file
# bac1 changed to centered value of bac1
# get the data within the bandwidth
dat_bandwidth_0.05 <- bac_data %>% filter(abs(bac1_ctd) <= 0.05)

# check being white
white_rdd <- dat_bandwidth_0.05 %>% lm(white ~ dui*bac1_ctd, data = .)

# use a customary function to calculate robust SEs
summaryR.lm(white_rdd, type = "hc1") 

# check being male
male_rdd <- dat_bandwidth_0.05 %>% lm(male ~ dui*bac1_ctd, data = .)
summaryR.lm(male_rdd, type = "hc1") 

# check accident
accident_rdd <- dat_bandwidth_0.05 %>% lm(acc ~ dui*bac1_ctd, data = .)
summaryR.lm(accident_rdd, type = "hc1") 

# check age
age_rdd <- dat_bandwidth_0.05 %>% lm(aged ~ dui*bac1_ctd, data = .)
summaryR.lm(age_rdd, type = "hc1") 
```

There is no evidence that the three demographic characteristics or accident rate were different between people who were just below and those just above the cutoff point. 

## Notes for future improvement

- Compute confidence intervals
- Incorporate logistic regression to estimate white, male, and accident
