---
title: "20221215_manipulation_testing"
author: "Yurun (Ellen) Ying"
date: "2022-12-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse);library(haven);library(here);library(rddensity);library(broom)

# input data
bac_data <- 
  here("summative/data/hansen_dwi.dta") %>% 
  read_dta() %>% 
  mutate(
    # code into drive under influence (dui) = 1 if bac1 >= 0.08
    dui = ifelse(bac1 >= 0.08, 1, 0),
    
    # create a quadratic term from bac1
    bac1_sq = bac1^2
    
    )

```

## Testing for density discontinuity at the cutoff point

We apply the estimator introduced by [Cattaneo, Jansson, and Ma (2020)](https://doi.org/10.1080/01621459.2019.1635480). To test whether there is a discontinuity of the probability density function at the cutoff point.

```{r}
# set cutoff
bac_cutoff <- 0.08

# get a vector of bac1
bac1 <- bac_data %>% pull(bac1)

# test for density discontinuity
bac_manipulation_test <- bac1 %>% rddensity(c = bac_cutoff) # set the cutoff point

# check the results
bac_manipulation_test %>% summary() 

# get the plot range
plot_min <- bac_cutoff - 3*bac_manipulation_test$h$left
plot_max <- bac_cutoff + 3*bac_manipulation_test$h$right

# plot the results
rdplotdensity(bac_manipulation_test, bac1,
              histBreaks = seq(plot_min, plot_max, by = 0.001) - 0.0005, # adjust the break to remove heaping
              xlabel = "Blood Alcohol Content (BAC)",
              ylabel = "Density",
              title = "Empirical and Estimated Density of Blood Alcohol Content",
              noPlot = TRUE
              ) 
```

We cannot reject the null hypothesis that there is no discontinuity in the pdf at the cutoff point. The result indicates that there is no manipulation at the cutoff point.